### 2024-10-21 manuscript.tex
Various improvements

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
%\setmainfont{Crimson Text}
\setmainfont{CrimsonText}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
%\setmonofont[Scale=0.8]{Space Mono}
\setmonofont[Scale=0.8]{SpaceMono}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald and Thorsten Mauritsen}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{12pt}{4pt}
%\renewcommand{\bibsection}{\section*{Publications}}
\parskip=6pt
\parindent=0pt

@@

\begin{center} \Large \sffamily\textbf{Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses}\\[0.4cm]
\large Peter Kuma$^\mathsf{1,}$*, Frida A.-M. Bender$^\mathsf{1}$, Adrian J. McDonald$^\mathsf{2}$ and Thorsten Mauritsen$^\mathsf{1}$\\[0.4cm]
\small
$^\mathsf{1}$Department of Meteorology (MISU) and Bolin Centre for Climate Research, Stockholm University, Stockholm SE-106 91, Sweden\\[0.2cm]
$^\mathsf{2}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, New Zealand\\[0.2cm]
*Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\large \today\\[0.4cm]

@@

Currently, a new generation of km-scale (also called `k-scale') resolution
global climate models is in development as the forthcoming phase of climate
modelling. One such model is a 5-km version of the Icosahedral Nonhydrostatic
Weather and Climate Model (ICON), developed jointly by the Deutscher
Wetterdienst (DWD) and Max-Planck-Institute for Meteorology (MPI-M). Because of
the high resolution, most parametrisations, such as those of convection and
clouds, can be avoided.

Previous studies have identified substantial large-scale biases in model clouds
over the Southern Ocean (SO), affecting sea surface temperature and the Earth's
albedo overall. Our aim is to quantify how well the high-resolution ICON model
is simulating clouds in this region, particularly in light of the fact that
subgrid-scale clouds are not parametrised in this model. This region is mostly
dominated by boundary layer clouds generated by shallow convection, and
these are problematic to observe by spaceborne lidar and radars, which are
affected by attenuation by overlapping and thick clouds and ground clutter,
respectively. Therefore, we choose to use a large set of ship-based
observations conducted with ceilometers and lidars on board of the Research
Vessel (RV) \textit{Polarstern} and other voyages. Altogether, we analyse over
2000 days (about 6 years) of data from 31 voyages and 1 sub-antarctic station
covering diverse longitudes of the SO. To achieve a like-for-like comparison
with the model, we use a ground-based lidar simulator called the Automatic
Lidar and Ceilometer Framework (ALCF). We contrast the results with the ECMWF
Reanalysis 5 (ERA5) and the Modern-Era Retrospective analysis for Research and
Applications, Version 2 (MERRA-2).

We show that the model underestimates the total cloud fraction (CF) by about 10\%,
with overestimation of cloud below 2 km and underestimation of cloud above 2
km. The reanalyses also underestimate the total CF by about 20\%.
ERA5 overestimates cloud below 1 km but underestimates near-surface cloud or
fog. In addition to lidar data, we compare radiosonde profiles acquired on the
RV \textit{Polarstern} voyages with ICON. Notably, the model exhibits smaller
natural variability than observations, and its lifting condensation level tends
to be higher. This might explain why cloud occurrence is peaking higher in the
model (at 500 m) than in observations (at the surface).

The results imply that the SO cloud biases are still a significant issue in a
km-scale resolution model, even though an improvement over the lower-resolution
reanalyses is notable. More effort is needed to improve model cloud simulations
in this fast-changing and understudied region. The advancement from convection
and cloud parametrisation to cloud-resolving models might not solve this bias
without an additional effort.

\section{Introduction}

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity.  Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPGPUs) is progressing to enable km-scale Earth
system models (ESMs) and coupled atmosphsere--ocean general circulation models
(AOGCMs) in research conditions today and operationally in the forthcoming
years. Therefore, it represents a natural advancement in climate modeling.
Global strom-resolving models (GSRMs) are emerging as a new front in the
development of high-resolution global climate models, with horizontal grid
resolutions of about 2--8 km \citep{satoh2019,stevens2019}. This is enough to
resolve mesoscale convective storms, but smaller-scale convective plumes and
smaller-scale cloud structure remain unresolved. At a 5-km scale,
non-hydrostatic processes also become important, and for this reason such
models are generally non-hydrostatic. The terms global cloud-resolving models
(GCRMs), or global convection-permitting/-resolving models (GCPMs) are also
sometimes used interchangeably with GSRMs but imply that clouds or convection
are resolved explicitly, which is not entirely true for GSRMs, as this would
require an even higher horizontal resolution \citep{satoh2019}.  Representative
of these efforts is the DYnamics of the Atmospheric general circulation Modeled
On Non-hydrostatic Domains (DYAMOND) project \citep{stevens2019,dyamond}, which
is an intercomparison of nine global storm-resolving models (GSRMs) on two
40-day time periods in summer (1 August -- 10 September 2016) and winter (20
January - 1 March 2020). A new one-year GSRM intercomparison is currently
proposed by \cite{takasuka2024}, with the hope of also evaluating the seasonal
cycle and large-scale circulation.  An alternative to using a computantionally
costly GSRM is to train an artificial neural network on GSRM output and use it
for subgrid-scale clouds, as done with the GSRM ICON by \cite{grundner2022};
\cite{grundner2023}.

nextGEMS is an EU-funded project \citep{nextgems} focused on the research and
development of GSRMs at multiple modelling centres and universities in Europe.
The project also develops GSRM versions of the Icosahedral Nonhydrostatic
Weather and Climate Model (ICON), the Integrated Forecasting System (IFS), and
their ocean components at eddy-resolving resolutions: ICON-O coupled with ICON
and Finite-Element/volumE Sea ice-Ocean Model (FESOM) and Nucleus for European
Modelling of the Ocean (NEMO) coupled with IFS.  The project has so far
produced ICON and IFS simulations in three cycles (Cycle 1--3) and pre-final
simulation, with a final production simulation planned by the end of the
project. nextGEMS is not the only project developing GSRMs. Other GSRMs (or
GSRM versions of climate models) currently in development include:
Convection-Permitting Simulations With the E3SM Global Atmosphere Model
[SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM; \cite{satoh2008}],
Unified Model (UM), eXperimental System for High-resolution modeling for
Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de Recherche Petite
Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], NASA Goddard Earth Observing System global
atmospheric model version 5 [GEOS5; \cite{putman2011}], Model for Prediction
Across Scales [MPAS; \cite{skamarock2012}], and System for Atmospheric Modeling
[SAM; \cite{khairoutdinov2003}].

Multiple cloud properties have an effect on the shortwave (SW) and longwave
(LW) radiation. On the first order, it is the total cloud fraction, cloud
phase, and the liquid and ice water path. These are influenced by the
atmopheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. On smaller orders, it is the cloud droplet size
distribution, ice crystal habit, cloud lifetime, and direct radiative
interaction with aerosols.  In the 6th phase of the Coupled Model
Intercomparison Project [CMIP6; \citep{eyring2016}], the cloud feedback has
increased relative to CMIP5 \citep{zelinka2020}, also being one of the main
reasons for the higher climate sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to reasons such as a lack of surface and in situ observations and
being a lower priority region for numerical weather prediction (NWP) and
climate model development because it is far away from populated areas of
interest.  Nevertheless, radiation biases and changes over an area of its size
have a very significant influence on the global climate, and the SO is an
important part of the global ocean conveyor belt.  Marine clouds have
disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The longitudinal symmetry of the SO
means that model cloud biases tend to be similar across longitudes.  Here, we
will conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW CRE
bias is still positive in eight analysed CMIP6 models analysed by
\cite{schuddeboom2021} over high-latitude SO, whereas over low-latitude SO it
tends to be more neutral or negative in some models. Too much absorbed SW
radiation over the SO was also identified in the GSRM SCREAM
\cite{caldwell2021}. Compensating biases are possible, such as the `too few too
bright' cloud bias, characterised by too small cloud fraction and too large
cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a resonable SW radiation balance by
reflecting too much SW radiation by cloudy sky but has too little cloudy sky
area overall. A study by \cite{konsta2022} showed that this type bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \citep{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to the Moderate Resolution Imaging
Spectroradiometer (MODIS)] and overestimates downwelling TOA SW radiation on
the order of 10 Wm$^\mathrm{-2}$ in  [relative to the Clouds and the Earth’s
Radiant Energy System (CERES)] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO.

@@

The main aim of our research is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst (DWD) and the Max-Planck-Institute
for Meteorology (MPI-M). Previous studies have identified substantial
large-scale biases in climate model clouds over the Southern Ocean (SO),
affecting sea surface temperature and the Earth’s albedo overall. Our aim is to
quantify how well the high-resolution ICON model is simulating clouds in this
region, particularly in light of the fact that subgrid-scale clouds are not
parameterised in this model. This region is mostly dominated by boundary layer
clouds generated by shallow convection, and these are problematic to observe by
spaceborne lidar and radars, which are affected by attenuation by overlapping
and thick clouds and ground clutter, respectively.  Specifically, the radar on
CloudSat and lidar on CALIPSO are affected by the abovementioned issues,
resulting in a strong underestimation of cloud occurrence below 2 km relative
to ground-based lidar observations \citep{mcerlich2021}.  This, in turn, can
lead to systematic biases in low clouds in climate models, which are frequently
evaluated against CloudSat--CALIPSO products. Reanalyses can also suffer from
cloud biases, as these are usually parametrised in their atmospheric component,
and also in regions where input observations are sparse.  This makes them a
problematic reference for clouds over the SO, and any biases relative to a
reanalysis should be interpreted with caution. Instead, we chose to use a large
set of ship-based observations conducted with ceilometers and lidars on board
of the RV \emph{Polarstern} and other voyages and stations as a reference for
the model evaluation.

Altogether, we analysed over 1500 days of data from 31 voyages and one
sub-Antarctic station covering diverse longitudes of the SO. To achieve a
like-for-like comparison with the model, we used a ground-based lidar simulator
called the Automatic Lidar and Ceilometer Framework (ALCF) \citep{kuma2021}. We
contrasted the results with the European Centre for Medium-Range Weather
Forecasts (ECMWF) Reanalysis 5 (ERA5) \citep{era5} and the Modern-Era
Retrospective analysis for Research and Applications, Version 2 (MERRA-2)
\citep{gelaro2017}.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
operating at a wavelenght of 910 nm and the Lufft CHM-15k operating at
1064 nm, described in detail below (Sections \ref{sec:cl51} and
\ref{sec:chm15k}). A ceilometer is a low-power near-infrared vertically
pointing lidar with the purpose of measuring the cloud base, but it also
measures the full vertical structure of clouds as long as the laser signal is
not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

Apart from lidar observations, radiosondes were launched on weather balloons at
regular synoptic times on the RV \emph{Polarstern}, MARCUS, NBP17024, TAN1702,
and TAN1802 voyages and campaigns, measuring pressure, temperature, relative
humidity, and global navigation satellite system coordinates. Derived
thermodynamic (virtual potential temperature, lifting condensation level, etc.)
and dynamic physical quantities (wind speed and direction) for the measured
vertical profiles were calculated with rstool \citep{rstool}. Surface
meteorological quantities were measured continuously by an onboard automatic
weather station or individual instruments.

>>>>

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage{microtype}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
\setmainfont{EB Garamond}
%\setmainfont{Crimson Text}
%\setmainfont{CrimsonText}[
%	Path=./fonts/,
%	Extension=.ttf,
%	UprightFont=*-Regular,
%	BoldFont=*-Bold,
%	ItalicFont=*-Italic,
%	BoldItalicFont=*-BoldItalic
%]
%\setmonofont[Scale=0.8]{Space Mono}
\setmonofont[Scale=0.8]{SpaceMono}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, …}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{13pt}{4pt}
%\renewcommand{\bibsection}{\section*{Publications}}
\parskip=6pt
\parindent=0pt

@@

\begin{center} \Large \sffamily\textbf{Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses}\\[0.4cm]
\large Peter Kuma$^\mathsf{1,}$*, Frida A.-M. Bender$^\mathsf{1}$, \ldots\\[0.4cm]
\small
$^\mathsf{1}$Department of Meteorology (MISU) and Bolin Centre for Climate Research, Stockholm University, Stockholm SE-106 91, Sweden\\[0.2cm]
*Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\large \today\\[0.4cm]

@@

Global storm-resolving models (GSRMs) are developed as the forthcoming phase of
climate modelling. Among them is a 5-km Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). Due to the resolution, parametrisations of convection and
clouds are avoided. Standard-resolution models have substantial cloud biases
over the Southern Ocean (SO), affecting radiation and sea surface temperature.
We quantify how the GSRM ICON simulates clouds over the SO and contrast it with
the ECMWF Reanalysis 5 (ERA5) and Modern-Era Retrospective analysis for
Research and Applications, Version 2 (MERRA-2). The SO is dominated by boundary
layer clouds, which often cannot be observed properly by speceborne instruments
due to overlapping clouds, attenuation, and ground clutter. Instead, we use
about 2400 days of ship-based lidar observations from 31 voyages and one
subantarctic station. We use a ground-based lidar simulator for a like-for-like
comparison. ICON underestimates the total cloud fraction (TCF) by about 10\%,
with overestimation of clouds below 2 km and underestimation of clouds above 2
km. The reanalyses underestimate TCF by about 20\%. ERA5 overestimates clouds
below 1 km but underestimates near-surface clouds or fog. Compared to
radiosonde observations, ICON exhibits smaller natural variability, and its
lifting condensation level is higher, which might explain why cloud occurrence
peaks higher in the model than in observations. SO cloud biases are still a
substantial issue in a GSRM, even though an improvement over the
lower-resolution reanalyses is notable. The transition from parametrisation to
explicitly resolved convection and cloud processes might not be enough to
address the model cloud biases.

\section{Introduction}
\label{sec:introduction}

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPGPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphsere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advancement in climate modeling.  Global strom-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models (GCRMs), or global
convection-permitting/-resolving models (GCPMs) are also sometimes used
interchangeably with GSRMs but imply that clouds or convection are resolved
explicitly, which is not entirely true for GSRMs, as this would require an even
higher horizontal resolution \citep{satoh2019}.  Representative of these
efforts is the DYnamics of the Atmospheric general circulation Modeled On
Non-hydrostatic Domains (DYAMOND) project \citep{stevens2019,dyamond}, which is
an intercomparison of nine global storm-resolving models (GSRMs) on two 40-day
time periods in summer (1 August -- 10 September 2016) and winter (20 January -
1 March 2020). A new one-year GSRM intercomparison is currently proposed by
\cite{takasuka2024}, with the hope of also evaluating the seasonal cycle and
large-scale circulation.  An alternative to using a computantionally costly
GSRM is to train an artificial neural network on GSRM output and use it for
subgrid-scale clouds, as done with the GSRM ICON by \cite{grundner2022};
\cite{grundner2023}.

nextGEMS is a European Union--funded project \citep{nextgems} focused on the research and
development of GSRMs at multiple modelling centres and universities in Europe.
The project also develops GSRM versions of the Icosahedral Nonhydrostatic
Weather and Climate Model (ICON), the Integrated Forecasting System (IFS), and
their ocean components at eddy-resolving resolutions: ICON-O coupled with ICON
and Finite-Element/volumE Sea ice-Ocean Model (FESOM) and Nucleus for European
Modelling of the Ocean (NEMO) coupled with IFS.  The project has so far
produced ICON and IFS simulations in three cycles (Cycle 1--3) and pre-final
simulation, with a final production simulation planned by the end of the
project. nextGEMS is not the only project developing GSRMs. Other GSRMs (or
GSRM versions of climate models) currently in development include:
Convection-Permitting Simulations With the E3SM Global Atmosphere Model
[SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM; \cite{satoh2008}],
Unified Model (UM), eXperimental System for High-resolution modeling for
Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de Recherche Petite
Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], the National Aeronautics and Space Administration
(NASA) Goddard Earth Observing System global atmospheric model version 5
[GEOS5; \cite{putman2011}], Model for Prediction Across Scales [MPAS;
\cite{skamarock2012}], and System for Atmospheric Modeling [SAM;
\cite{khairoutdinov2003}].

Multiple cloud properties have an effect on the shortwave (SW) and longwave
(LW) radiation. On the first order, it is the total cloud fraction, cloud
phase, and the liquid and ice water path. These are influenced by the
atmopheric thermodynamics, convection and circulation, and indirect and direct
effects of aerosols. On smaller orders, it is the cloud droplet size
distribution, ice crystal habit, cloud lifetime, and direct radiative
interaction with aerosols.  In the 6$^\mathrm{th}$ phase of the Coupled Model
Intercomparison Project [CMIP6; \citep{eyring2016}], the cloud feedback has
increased relative to CMIP5 \citep{zelinka2020}, also being one of the main
reasons for the higher climate sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to reasons such as a lack of surface and in situ observations and
being a lower priority region for numerical weather prediction (NWP) and
climate model development because it is far away from populated areas of
interest.  Nevertheless, radiation biases and changes over an area of its size
have a very substantial influence on the global climate, and the SO is an
important part of the global ocean conveyor belt.  Marine clouds have
disproportionate effect on top of atmosphere (TOA) SW radiation due to the
relatively low albedo of the sea surface.  The relative longitudinal symmetry
of the SO means that model cloud biases tend to be similar across longitudes.
Here, we conventionally refer to the SO as ocean regions south of 40°S,
low-latitude SO as 40--55°S and high-latitude SO as south of 55°S.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW CRE
bias is still positive in eight analysed CMIP6 models analysed by
\cite{schuddeboom2021} over high-latitude SO, whereas over low-latitude SO it
tends to be more neutral or negative in some models. Too much absorbed SW
radiation over the SO was also identified in the GSRM SCREAM
\cite{caldwell2021}. Compensating biases are possible, such as the `too few too
bright' cloud bias, characterised by too small cloud fraction and too large
cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models. That is, a model maintains a resonable SW radiation balance by
reflecting too much SW radiation by cloudy sky but has too little cloudy sky
area overall. A study by \cite{konsta2022} showed that this type bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \citep{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to the Moderate Resolution Imaging
Spectroradiometer (MODIS)] and overestimates downwelling TOA SW radiation on
the order of 10 Wm$^\mathrm{-2}$ in  [relative to the Clouds and the Earth’s
Radiant Energy System (CERES)] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulation over a
SO region adjacent to Tasmania agaist the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations. They found the ICON cloud optical
thickness underestimated relative to Himawari‐8 satellite observations, but also
large differences in cloud top phase.

@@

The main aim of this study is to evaluate the GSRM version of ICON, developed
jointly by nextGEMS, Deutscher Wetterdienst (DWD), Max-Planck-Institute for
Meteorology (MPI-M), Deutsches Klimarechenzentrum (DKRZ), Karlsruhe Institute
of Technology (KIT), and the Center for Climate Systems Modeling (C2SM).
Previous studies have identified substantial large-scale biases in climate
model clouds over the SO, affecting sea surface temperature and the Earth’s
albedo. Our aim is to quantify how well the GSRM ICON is simulating clouds in
this region, particularly in light of the fact that subgrid-scale clouds and
convection are not parameterised in this model. This region is mostly dominated
by boundary layer clouds generated by shallow convection, and these are
problematic to observe by spaceborne lidars and radars, which are affected by
attenuation by overlapping and thick clouds and ground clutter, respectively.
Specifically, the radar on CloudSat and lidar on CALIPSO (neither of them
operational any more) are affected by the abovementioned issues, resulting in a
strong underestimation of cloud occurrence below 2 km relative to ground-based
lidar observations \citep{mcerlich2021}.  This, in turn, can lead to systematic
biases in low clouds in climate models, which are frequently evaluated against
CloudSat--CALIPSO products. Reanalyses can also suffer from cloud biases, as
these are usually parametrised in their atmospheric component, and also in
regions where input observations are sparse.  This makes them a problematic
reference for clouds over the SO, and any biases relative to a reanalysis
should be interpreted with caution. Instead, we chose to use a large set of
ship-based observations conducted with ceilometers and lidars on board of the
RV \emph{Polarstern} and other voyages and stations as a reference for the
model evaluation.

Altogether, we analysed about 2400 days of data from 31 voyages and one
sub-Antarctic station covering diverse longitudes and latitudes of the SO. To
achieve a like-for-like comparison with the model, we used a ground-based lidar
simulator called the Automatic Lidar and Ceilometer Framework (ALCF)
\citep{kuma2021}. We contrasted the results with the European Centre for
Medium-Range Weather Forecasts (ECMWF) Reanalysis 5 (ERA5) \citep{era5} and the
Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) \citep{gelaro2017}.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
 perating at a wavelenght of 910 nm, Vaisala CT25K operating at 905 nm, and
Lufft CHM 15k operating at 1064 nm, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar with the purpose of measuring the cloud base, but it
also measures the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

Apart from lidar observations, radiosondes were launched on weather balloons at
regular synoptic times on the RV \emph{Polarstern}, MARCUS, NBP17024, TAN1702,
and TAN1802 voyages and campaigns, measuring pressure, temperature, relative
humidity, and the global navigation satellite system coordinates. Derived
thermodynamic (virtual potential temperature, lifting condensation level, etc.)
and dynamic physical quantities (wind speed and direction) for the measured
vertical profiles were calculated with rstool \citep{rstool}. Surface
meteorological quantities were measured continuously by an onboard automatic
weather station or individual instruments.

\begin{table}[p!]
% TODO: Add sampling intervals.
\caption{
An overview of the analysed voyages, campaigns and stations. Start, end, and
the number of days (UTC; inclusive) refer to the time period when the vessel
was south of 40°S.  Abbreviations: ceilometer (ceil.), Australia (AU), New
Zealand (NZ), South America (SA), Atlantic Ocean (AO), and Africa (AF). The
number of days is rounded to the nearest integer. CL51/31 indicates CL51
configured to emulate CL31.
}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
\textbf{Name} & \textbf{Vessel or station} & \textbf{Ceil.} & \textbf{Region} & \textbf{Start} & \textbf{End} & \textbf{Days}\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124 \\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CT25K   & AU       & 2017-10-29 & 2018-03-26 & 149 \\
MICRE    & Macquarie Is. station         & CT25K   & AU/NZ    & 2016-04-03 & 2018-03-14 & 710 \\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}
